\documentclass[
11pt,%
tightenlines,%
twoside,%
onecolumn,%
nobibnotes,%
nofootinbib,%
superscriptaddress,%
noshowpacs,%
centertags]%
{revtex4-2}
\usepackage{ljm}
\setcaptionwidth{0.93\textwidth}

% --- ГРАФИКА (WINDOWS) ---
\usepackage{epstopdf} % Критично для Windows: конвертирует EPS на лету
\graphicspath{{figures/}} 

% --- СТИЛЬ ЖУРНАЛА ---
\usepackage{ljm} % Файл должен лежать рядом!

% --- МАТЕМАТИКА И ТАБЛИЦЫ ---
\usepackage{mathtools}
\usepackage{colortbl}
\usepackage{algorithm2e}
% \usepackage{multirow}
\usepackage{makecell}  %new line in one table cell

% --- TIKZ (ВНИМАНИЕ: Для журнала в итоге нужны EPS файлы) ---
\usepackage{pgf}
\usepackage{tikz}
\usepackage{nicematrix}
\usetikzlibrary{arrows, automata, fit, calc, positioning}

% --- ДОПОЛНИТЕЛЬНО ---
% \usepackage{soul}
\usepackage[makeroom]{cancel}
\newcommand{\mycomment}[1]{}

\usepackage[caption=false]{subfig}
\usepackage{wrapfig}

\begin{document}
%
\title{GLL-based Context-Free Path Querying for Neo4j}

% --- АВТОРЫ (Формат RevTeX) ---

\author{Vadim Abzalov}
\email[E-mail: ]{vadim.i.abzalov@gmail.com}
\affiliation{Saint Petersburg State University, St. Petersburg, Russia}

\author{Vlada Pogozhelskaya}
\email[E-mail: ]{vvpogozhelskaia@edu.hse.ru}
\affiliation{National Research University Higher School of Economics, St. Petersburg, Russia}

\author{Vladimir Kutuev}
\email[E-mail: ]{v.kutuev@spbu.ru}
\affiliation{Saint Petersburg State University, St. Petersburg, Russia}

\author{Olga Bachishche}
\email[E-mail: ]{bachisheo@gmail.com}
\affiliation{ITMO University, St. Petersburg, Russia}

\author{Semyon Grigorev}
\email[E-mail: ]{s.v.grigoriev@spbu.ru}
\affiliation{Saint Petersburg State University, St. Petersburg, Russia}


% --- АННОТАЦИЯ (Должна быть ДО maketitle) ---
\begin{abstract}
We propose a GLL-based \textit{context-free path querying} algorithm that handles queries in Extended Backus-Naur Form (EBNF) using Recursive State Machines (RSM). Utilization of EBNF allows one to combine traditional regular expressions and mutually recursive patterns in constraints natively. The proposed algorithm solves both the \textit{reachability-only} and the \textit{all-paths} problems for the \textit{all-pairs} and the \textit{multiple sources} cases. The evaluation on real-world graphs demonstrates that the utilization of RSMs increases the performance of query evaluation. Being implemented as a stored procedure for Neo4j, our solution demonstrates better performance than a similar solution for RedisGraph. The performance of our solution on regular path queries is comparable to the performance of the native Neo4j solution, and in some cases, our solution requires significantly less memory.
\end{abstract}

% --- КЛЮЧЕВЫЕ СЛОВА ---
\keywords{Graph database, graph querying, context-free path querying, CFPQ, reachability problem, all-paths problem, generalized LL, GLL}

% --- ГЕНЕРАЦИЯ ЗАГОЛОВКА ---
\maketitle
\renewcommand{\figurename}{Fig.} 
% --- MAIN CONTENT ---
\input{sections/introduction}
\input{sections/preliminaries}
\input{sections/rsm}
\input{sections/gll}
\input{sections/dataset}
\input{sections/bnf_vs_ebnf}
\input{sections/cfpq_for_neo4j}
\input{sections/related_work}
\input{sections/conclusion}

\section*{ACKNOWLEDGMENTS}
We thank Adrian Johnstone for pointing out the Generalized LL algorithm in our discussion at Parsing@SLE--2013, which motivated the development of the presented solution. We thank George Fletcher for the discussion of the evaluation of different CFPQ algorithms for Neo4j.

\section*{FUNDING}
The study was funded by the St. Petersburg State University (Grant No. 116636233). 

\section*{CONFLICT OF INTEREST}
The authors of this work declare that they have no conflicts of interest.

%
% ---- Bibliography ----
%
% The bibliography style 'APS (American Physical Society)' is a standad for RevTex.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{apsrev4-2}
\bibliography{gll}
%
\end{document}